To be FAIR: Theory Specification Needs an Update

### 2025-02-25 manuscript.Rmd
Update manuscript.Rmd

FAIR theory imposes no restrictions on the manner in which theories are derived and implemented;
rather, it increases the fidelity and ease with which they are communicated.
Thus, FAIR theory does not require theories to be formal.
At the same time, not every formal theory is FAIR.
The FAIR principles pertain to theories' metadata documentation and sharing in digital archives, with the aim of enhancing their reusability and extensibility.
They can be applied to theories representated in natural language,
as well as formal theories represented using mathematical notation, algorithmic pseudo code, or a set of logical clauses.
Thus, for example, "grounded theory", derived from qualitative research,
can be represented as a FAIR theory if it is represented as plain-text propositions and archived in a FAIR repository with appropriate metadata.
Conversely, a formal theory is not FAIR if it is confined to in a journal article without any key words to identify it as a theory paper (lacking Findability), represented as a bitmap image (limiting Accessibility and Interoperability), and behind a paywall (limiting Reusability).
FAIR theory is thus consistent with, but does not require, formalization (also see *Accessibility*).

@@

```
Given a version number MAJOR.MINOR.PATCH, increment the:

MAJOR version when you make backwards incompatible changes, i.e., the
      theory now contains empirical statements that are at odds with
      a previous version
MINOR version when you expand the set of empirical statements in a
      backward compatible manner (i.e., the previous version is
      subsumed within the new version)
PATCH version when you make backward compatible bug fixes, cosmetic
      changes, fix spelling errors, or add clarifications
```

>>>>

FAIR theory imposes no restrictions on the manner in which theories are derived and implemented;
rather, it increases the fidelity and ease with which they are communicated.
Thus, FAIR theory does not require theories to be formal.
At the same time, not every formal theory is FAIR.
The FAIR principles pertain to theories' metadata documentation and sharing in digital archives, with the aim of enhancing their reusability and extensibility.
They can be applied to theories representated in natural language,
as well as formal theories represented using mathematical notation, algorithmic pseudo code, or a set of logical clauses.
Thus, for example, "grounded theory", derived from qualitative research,
can be represented as a FAIR theory if it is represented as plain-text propositions and archived in a FAIR repository with appropriate metadata.
Conversely, a formal theory is not FAIR if it is confined to a publication in a journal article without any key words to identify it as a theory paper (lacking Findability), represented merely as a bitmap image (limiting Accessibility and Interoperability), or behind a paywall (limiting Reusability).
FAIR theory is thus consistent with, but does not require, formalization (also see *Accessibility*).

@@

Given a version number in the format MAJOR.MINOR.PATCH (where MAJOR, MINOR, and PATCH are placeholders for positive integer numbers including zero), increment the:

- MAJOR version when you commit backwards incompatible changes, i.e., the
      theory now contains empirical statements that are at odds with
      a previous version of the theory
- MINOR version when you expand the set of empirical statements in a
      backward compatible manner (i.e., the previous version is
      subsumed within the new version)
- PATCH version when you make backward compatible bug fixes, cosmetic
      changes, fix spelling errors, or add clarifications
